---
title: '2'
output: html_document
---
---
title: "R Notebook"
output: html_notebook
---
Introduction:
Wisconsin Breast Cancer Data Alanysis

Loading libraries:
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(ggfortify)
library(fastICA)
library(cluster)
library(caret)
library(funModeling)
library(randomForest)
library(mlbench)
library(glmnet)
library(relaimpo)
library(corrplot)
library(ROCR)
library(pROC)
library(MASS)
library(gridExtra)
library(ISLR)
library(tree)
library(gbm)
library(rpart)
library(e1071)
```

Data cleaning:
```{r}
data(BreastCancer)
data <- BreastCancer 
data <- na.omit(data)
names(data)[11]<-"Diagnosis"
names(data)[10]<-"Mitosis"
data$Id <- as.numeric(data$Id)
data$Cl.thickness <- as.numeric(data$Cl.thickness)
data$Cell.size <- as.numeric(data$Cell.size)
data$Cell.shape <- as.numeric(data$Cell.shape)
data$Marg.adhesion <- as.numeric(data$Marg.adhesion)
data$Epith.c.size <- as.numeric(data$Epith.c.size)
data$Bare.nuclei <- as.numeric(data$Bare.nuclei)
data$Bl.cromatin <- as.numeric(data$Bl.cromatin)
data$Normal.nucleoli <- as.numeric(data$Normal.nucleoli)
data$Mitosis <- as.numeric(data$Mitosis)
glimpse(data)
```

Proportion of benign vs malignant diagnosis
```{r}
(round(prop.table(table(data$Diagnosis)), 2))
plot(data$Diagnosis)
```

Correlation of variables:
```{r}
correlation_matrix = cor(data[,1:10])
corrplot(correlation_matrix, order = "hclust", method = "square")
```

Variable Distribution:
```{r}
dataclust <- data[2:10]

pdf("distv2.pdf", height = 10, width = 15)
plot_num(dataclust)
dev.off()
```


Dendrogram:
```{r}
hierarchical_dist <- dist(dataclust, method = "euclidean")
tree <- hclust(hierarchical_dist, method="average")

pdf("dendrogramv2.pdf", height =15, width = 50)
plot(tree)
dev.off() 
```


PCA:
```{r}

pca2 <- data.matrix(dataclust)

Center <- apply(pca2, 2, function(x) x - mean(x))
Covariance <- cov(Center)
Eigen_value <- eigen(Covariance)$value

Eigen_vector <- eigen(Covariance)$vector
PC <- as.data.frame(data.matrix(Center) %*% Eigen_vector)

ggplot(PC, aes(PC[,1], PC[,2])) + geom_point(aes(PC[,1], PC[,2]))
round(cumsum(Eigen_value)/sum(Eigen_value) * 100, digits = 2)

autoplot(prcomp(pca2))
```


ICA:
```{r}
a <- fastICA(dataclust, 7, alg.typ = "parallel", fun = "logcosh", alpha = 1,
             method = "R", row.norm = FALSE, maxit = 200,
             tol = 0.0001, verbose = TRUE)
heatmap(a$S)
```

Modeling:

Split into test/training sets:
```{r}
dt = sort(sample(nrow(data), nrow(data)*.75))
trainset = data[dt,] 
testset = data[-dt,]
prop.table(table(trainset$Diagnosis))
```

Feature Selection:

RFE
```{r}
control = rfeControl(functions = caretFuncs, number = 2)
results = rfe(data[,1:10], data[,11], sizes = c(2,5,9,11), rfeControl = control, method = "svmRadial")
results
results$variables
```

Random Forests:
```{r}
train <- floor(0.75 * nrow(data))
train1 <- sample(seq_len(nrow(data)), size = train)
train_class <- data[train1, ]
test_class <- data[-train1, ]
dim(train_class)
dim(test_class)

rfmodel = randomForest(Diagnosis ~ Cl.thickness + Cell.size + Cell.shape + Marg.adhesion + Epith.c.size + Bare.nuclei + Bl.cromatin + Normal.nucleoli +  Mitosis, data=train_class,  importance = TRUE, oob.times = 15, confusion = TRUE)
importance(rfmodel)
```

Simulated Annealing:
```{r}
ctrl <- safsControl(functions = rfSA,
                        method = "repeatedcv",
                        repeats = 3,
                        improve = 5) 

set.seed(100)
sa <- safs(x=data[, c(1:10)], 
               y=data[, 11],
               safsControl = ctrl)


print(sa$optVariables)
```

Top 5 Variables: Cl.thickness, Cell.size, Cell.shape, Epith.c.size, Bl.cromatin

Classification:

LDA:
```{r}
Clthickness <- ggplot(data = data, aes(x = Cl.thickness, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10)  +
  theme_bw()
Cellsize <- ggplot(data = data, aes(x = Cell.size, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Cellshape <- ggplot(data = data, aes(x = Cell.shape, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Margadhesion <- ggplot(data = data, aes(x = Marg.adhesion, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Epithcsize <- ggplot(data = data, aes(x = Epith.c.size, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Barenuclei <- ggplot(data = data, aes(x = Bare.nuclei, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Blcromatin <- ggplot(data = data, aes(x = Bl.cromatin, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Normalnucleoli <- ggplot(data = data, aes(x = Normal.nucleoli, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Mitosis1 <- ggplot(data = data, aes(x = Mitosis, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
grid.arrange(Clthickness, Cellsize, Cellshape, Margadhesion, Epithcsize, Barenuclei, Blcromatin, Normalnucleoli, Mitosis1)
```

LDA of top 5 variables: Graphs
```{r}
Clthickness <- ggplot(data = data, aes(x = Cl.thickness, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10)  +
  theme_bw()
Cellsize <- ggplot(data = data, aes(x = Cell.size, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Cellshape <- ggplot(data = data, aes(x = Cell.shape, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Epithcsize <- ggplot(data = data, aes(x = Epith.c.size, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()
Blcromatin <- ggplot(data = data, aes(x = Bl.cromatin, fill = Diagnosis)) + 
  geom_histogram(position="identity", alpha=0.5, bins= 10) +
  theme_bw()

grid.arrange(Clthickness, Cellsize, Cellshape, Epithcsize, Blcromatin)
```

LDA Data Top 5 Variables:
```{r}

LDA <- lda(Diagnosis~ Cl.thickness+ Cell.size+ Cell.shape+ Epith.c.size+ Bl.cromatin, 
           data= train_class, cv= T)
LDA
```

LDA Model Test:
```{r}
LDA_predict <- predict(LDA, newdata=test_class)
cm1 <- confusionMatrix(LDA_predict$class, reference = test_class$Diagnosis)
cm1

fourfoldplot(cm1$table)
```

```{r}
# save the predictions in a new variable
predictions <- as.data.frame(LDA_predict$posterior) %>% 
  rownames_to_column("idx")

head(predictions)

test_classifier <- test_class %>% 
  rownames_to_column("idx")

predictions_actual <- full_join(predictions,test_classifier, by = "idx" )

# choose the two classes we want to compare, setosa and versicolor
set_vers_true_labels <- predictions_actual %>% 
  filter(Diagnosis %in% c("benign", "malignant")) %>% 
  mutate(Diagnosis = as.character(Diagnosis)) 
  
#make dataframe of the prediction and the label
pred_label <- data.frame(prediction = set_vers_true_labels$malignant,
                         label = set_vers_true_labels$Diagnosis)

#ggplot(pred_label, aes(x = 1:24, y = prediction, color = label))+
  #geom_point()

pred <- prediction(set_vers_true_labels$malignant, set_vers_true_labels$Diagnosis, 
label.ordering = c("benign", "malignant")) 

perf <- performance(pred,"tpr","fpr")
plot(perf)
```


Logistic Regression:
```{r}
train_classifier_log <- train_class[c(which(train_class$Diagnosis == "benign"),
                                           which(train_class$Diagnosis == "malignant")),]
test_classifier_log <- test_class[c(which(test_class$Diagnosis == "benign"), 
                                         which(test_class$Diagnosis == "malignant")),]

train_classifier_log$Diagnosis <- factor(train_classifier_log$Diagnosis)
test_classifier_log$Diagnosis <- factor(test_classifier_log$Diagnosis)

ctrl1 <- trainControl(method = "repeatedcv", repeats = 15,classProbs = T,
                     savePredictions = T)

logistic_regression <- train(Diagnosis~ Cl.thickness+ Cell.size+ Cell.shape+ Epith.c.size+ Bl.cromatin, data = train_classifier_log, method = "glm", family= "binomial", trControl = ctrl1)
logistic_regression
summary(logistic_regression)
```

ROC Curve:
```{r}
plot(x = roc(predictor = logistic_regression$pred$benign,
             response = logistic_regression$pred$obs)$specificities, 
     y = roc(predictor = logistic_regression$pred$benign, 
             response = logistic_regression$pred$obs)$sensitivities,
     col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity",
     xlab = "Specificity")
legend("bottomright", legend = paste("Benign vs Malignant --", 
                                     roc(predictor = logistic_regression$pred$benign,
                                         response = logistic_regression$pred$obs)$auc
, sep = ""), col = c("blue"), fill = c("blue"))
```

Logistic Regression Model Test:
```{r}
logistic_regression_predict_class <- predict(logistic_regression, 
                                             newdata = test_classifier_log)

#confusion matrix
cm2 <-confusionMatrix(logistic_regression_predict_class, 
                reference = test_classifier_log$Diagnosis)
cm2
fourfoldplot(cm2$table)
```

Decision Tree:
```{r}
model_dtree<- rpart(Diagnosis ~ Cl.thickness + Cell.size + Cell.shape + Epith.c.size + Bl.cromatin, data = train_class)       
preds_dtree <- predict(model_dtree, newdata = test_class, type = "class")
plot(preds_dtree, main="Decision tree created using rpart")
predict_dtree <- table(preds_dtree, test_class$Diagnosis)
cm3 <- confusionMatrix(predict_dtree)
cm3
fourfoldplot(cm3$table)
```

Naive Bayes
```{r}
model_naive<- naiveBayes(Diagnosis ~ Cl.thickness + Cell.size + Cell.shape + Epith.c.size + Bl.cromatin, data = train_class)   
preds_naive <- predict(model_naive, newdata = test_class)        
test_naive <- table(preds_naive, test_class$Diagnosis)

cm4 <- confusionMatrix(test_naive) 
cm4
fourfoldplot(cm4$table)
plot(preds_naive, main="Naive Bayes")
```

SVM
```{r}
control = trainControl(method = "repeatedcv", repeats = 5, classProbs = T, savePredictions = T)

svm = train(Diagnosis ~ Cl.thickness + Cell.size + Cell.shape + Epith.c.size + Bl.cromatin,  data = train_class, method = "svmLinear", tuneLength = 10, trControl = control)

svm
```

ROC 
```{r}
roc(predictor = svm$pred$malignant, response = svm$pred$obs)$auc

plot(x = roc(predictor = svm$pred$malignant, response = svm$pred$obs)$specificities, y = roc(predictor = svm$pred$malignant, response = svm$pred$obs)$sensitivities, col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity", xlab = "Specificity") 
legend("bottomright", legend = paste("Benign vs Malignant --", 
                                     roc(predictor = svm$pred$benign,
                                         response = svm$pred$obs)$auc))
```

SVM Test:
```{r}
svm_test = predict(svm, newdata = test_class)
cm5 <- confusionMatrix(svm_test, reference = test_class$Diagnosis)
cm5
fourfoldplot(cm5$table)
```

SVM Radial Kernel:
```{r}
control = trainControl(method = "repeatedcv", repeats = 5, classProbs = T, savePredictions = T)

svmr = train(Diagnosis ~ Cl.thickness + Cell.size + Cell.shape + Epith.c.size + Bl.cromatin,  data = train_class, method = "svmRadial", tuneLength = 10, trControl = control)

svmr
```

ROC:
```{r}
roc(predictor = svm$pred$malignant, response = svm$pred$obs)$auc

plot(x = roc(predictor = svm$pred$malignant, response = svm$pred$obs)$specificities, y = roc(predictor = svm$pred$malignant, response = svm$pred$obs)$sensitivities, col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity", xlab = "Specificity")
legend("bottomright", legend = paste("Benign vs Malignant --", 
                                     roc(predictor = svmr$pred$benign,
                                         response = svmr$pred$obs)$auc))
```

SVM Radial Test:
```{r}
svmr_test = predict(svmr, newdata = test_class)
confusionMatrix(svmr_test, reference = test_class$Diagnosis)
```

